\mutagen: Reliable Coverage-Guided, Property-Based Testing using Exhaustive Well-Typed Mutations

### 2021-03-31 papers/oopsla-21/main.tex
WIP

All the experiments were done in a dedicated workstation with an Intel Core
i7-8700 CPU running at 3.20GHz, and equipped with 32Gb of RAM.
%
We ran each experiment 30 times except for the ones involving the bugs on the
WebAssembly interpreter, which were run 10 times.
%
From there, we followed the same approach taken by
\citeauthor{lampropoulos2019coverage} and collected the Mean-Time-To-Failure
(MTTF) of each bug, i.e., how quickly a bug can be found in wall time.
%
In all cases, we used a one-hour timeout to stop the execution of both tools if
they have not yet found a counterexample.

@@

\paragraph{Grammar-based Fuzzing}

\paragraph{Random Data Generation}

%% Don't forget to cite "Deriving good generators ..."

@@

>>>>

All the experiments were performed in a dedicated workstation with an Intel Core
i7-8700 CPU running at 3.20GHz, and equipped with 32Gb of RAM.
%
We ran each experiment 30 times except for the ones involving the bugs on the
WebAssembly interpreter, which were run 10 times.
%
From there, we followed the same approach taken by
\citeauthor{lampropoulos2019coverage} and collected the Mean-Time-To-Failure
(MTTF) of each bug, i.e., how quickly a bug can be found in wall time.
%
In all cases, we used a one-hour timeout to stop the execution of both tools if
they have not yet found a counterexample.

@@

There exists a vast literature on fuzzing and property-based testing.
%
In this section we focus on three main topics:
%
automated random data generation using static information,
%
fuzzing using coverage information to guide mutations, and
%
exhaustive bounded testing tools, which inspired the design of \mutagen.

\paragraph{Automated Random Data Generation}

Obtaining good random generators automatically from static information (e.g.,
grammars or datatype definitions) has been tackled from different angles.

% DRAGEN and DRAGEN2
In Haskell, DRAGEN \cite{dragen} is a meta-programming tool that synthezises
random generators from datatypes definitions, using stochastic models to predict
and optimize their distribution based on a target distribution set by the user.
%
DRAGEN2 \cite{dragen2} extends this idea adding support for generating richer
random data by extracting additional static information like library APIs and
function input patterns from the codebase.
%
% Bolztmann-brain
Similarly, \cite{bendkowski} developed a polinomial tuning mechanism based on
Boltzmann samplers \cite{boltzmann} that synthesizes generators with
approximate-size distributions for combinatorial structures like lists or trees.
%

Automatically deriving random generators is substantially more complicated when
the generated data must satisfy by (often sparse) preconditions.
%
% Generating Constrained Random Data with Uniform Distribution
%
\cite{claessen2014} developed an algorithm for generating inputs constrained by
boolean preconditions with almost-uniform distribution.
%
% Luck
%
Later, \cite{luck} extends this approach by adding a limited form of constraint
solving controllable by the user in a domain-specific language called
\emph{Luck}.
%
% Deriving good generators
%
Recently, \cite{generatinggood} proposed a derivation mechanism to obtain
constrained random generators directly by defining their structure
using inductively defined relations in Coq.

We consider all these generational approaches to be orthogonal to the ideas
behind \mutagen.
%
In principle, our tool is tailored to improve the performance of poor
automatically derived generators.
%
However, more specialized generators can by used directly by \mutagen, and
combining them with well-typed mutations using perhaps a hybrid technique is an
appealing idea that we aim to address in the future.

\paragraph{Coverage-guided Fuzzing}

% AFL

% AFLfast

% CollAFL

% Criterion

% CSmith, XSmith

% QuickFuzz

@@

A popular category of property-based testing tools does not rely on randomness.
%
Instead, all possible inputs can be enumerated and tested from smaller to larger
up to a certain size bound.

%
% Feat
%
\emph{Feat} \cite{feat} formalizes the notion of \emph{functional enumerations},
For any algebraic type, it synthesizes a bijection between a finite prefix of
the natural numbers and set of increasingly growing values of the input type.
%
Later, this bijection can be traversed exhaustively or, more interestingly,
randomly accesed.
%
This allows the user to easily generate random values in a uniform basis simply
by sampling natural numbers.
%
However, values are enumerated based only on their type definition, so this
technique is not suitable for testing properties with sparse preconditions
expressed elsewhere.

% SmallCheck and LazySmallCheck
\emph{SmallCheck} \cite{smallcheck} is a Haskell tool that follows this idea.
%
\emph{SmallCheck} is an exhaustive testing tool for Haskell that progressively
executes the testing properties against all possible inputs values of within
a certain size bound.
%
Later, \emph{LazySmallCheck} is a lazy variant of \emph{SmallCheck} that uses
lazy evaluation to automatically prune the search space by detecting input
subexpressions that are not evaluated by the testing properties, and hence do
not need to be populated exhaustively with concrete values.

% Korat
On the object-oriented realm, \emph{Korat} \cite{korat} is a Java tool that uses
method specification predicates to automatically generate all (nonisomorphic)
test cases up to a given small size.
%
Similar to \emph{LazySmallCheck}, \emph{Korat} uses method preconditions to
prune the search space.
%
However, in \emph{Korat}, pruning is done by analyzing the execution trace of
the precondition predicate instead of by keeping track of unused subexpressions.

Being exhaustive, these approaches can only be used to enumerate and test a
relatively little space of small inputs.
%
In this work we use exhaustiveness as a way to reliably schedule different
mutants.
%
In contrast to fully-exhaustive testing tools, \mutagen relies initially on
randomness to find interesting mutable candidates.
%
Later, enumerating mutations for a given interesting candidate is much cheaper
than enumerating \emph{all posible values}.
%
Hence, our tool can be used to produce much larger test cases before their size
becomes computationally too restrictive.
